\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-11-26 papers/jss/article.Rnw
s/protocol buffers/Protocol Buffers/
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@911 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

One example of a system that supports Protocol Buffers to interact
with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework
for embedded statistical computation and reproducible research based
on \proglang{R} and \LaTeX. It exposes a HTTP(S) API to access and
manipulate \proglang{R} objects and execute remote \proglang{R}
function calls. Clients do not need to understand or generate any
\proglang{R} code: HTTP requests are automatically mapped to function
calls, and arguments/return values can be posted/retrieved using
several data interchange formats, such as Protocol Buffers.  OpenCPU
uses the \code{rexp.proto} descriptor and the \code{serialize\_pb} and
\code{unserialize\_pb} functions described in
Section~\ref{sec:evaluation} to convert between \proglang{R} objects
and protocol buffer messages.

@@

\begin{verbatim}
  GET https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb
\end{verbatim}
The postfix \code{/pb} in the URL tells the server to send this
object in the form of a protocol buffer message.
% Alternative formats include \code{/json}, \code{/csv}, \code{/rds} and others.
If the request
is successful, OpenCPU returns the serialized object with HTTP status
code 200 and HTTP response header \code{Content-Type: application/x-protobuf}.
The latter is the conventional MIME type that formally notifies the client to
interpret the response as a protocol buffer.

Because both HTTP and Protocol Buffers have libraries available for many
languages, clients can be implemented in just a few lines of code. Below
is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a protocol buffer message from OpenCPU.
In \proglang{R}, we use the HTTP client from
the \code{httr} package \citep{httr}. In this example we
download a data set which is part of the base \proglang{R} distribution, so we can
verify that the object was transferred without loss of information.

@@

This generates Python module called \code{rexp\_pb2.py}, containing
both the descriptor information as well as methods to read and
manipulate the \proglang{R} object message. We use the
HTTP client from the \code{urllib2} module in our example to retrieve the
encoded protocol buffer from the remote server then parse and print it
from Python.

@@

The previous example used a simple \code{HTTP GET} method to retrieve
an \proglang{R} object from a remote service (OpenCPU) encoded as a
protocol buffer.
In many cases simple \code{HTTP GET} methods are insufficient, and a
more complete RPC system may need to create compact protocol buffers
for each request to send to the remote server in addition to parsing
the response protocol buffers.

The OpenCPU framework allows us to do arbitrary \proglang{R} function
calls from within any programming language by encoding the arguments
in the request protocol buffer.  The following example \proglang{R}
client code performs the remote function call \code{stats::rnorm(n=42,
mean=100)}. The function arguments (in this case \code{n} and
\code{mean}) as well as the return value (a vector with 42 random
numbers) are transferred using protocol buffer messages. RPC in OpenCPU
works like the \code{do.call} function in \proglang{R}, hence all
arguments are contained within a list.

>>>>

One example of a system that supports Protocol Buffers to interact
with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework
for embedded statistical computation and reproducible research based
on \proglang{R} and \LaTeX. It exposes a HTTP(S) API to access and
manipulate \proglang{R} objects and execute remote \proglang{R}
function calls. Clients do not need to understand or generate any
\proglang{R} code: HTTP requests are automatically mapped to function
calls, and arguments/return values can be posted/retrieved using
several data interchange formats, such as Protocol Buffers.  OpenCPU
uses the \code{rexp.proto} descriptor and the \code{serialize\_pb} and
\code{unserialize\_pb} functions described in
Section~\ref{sec:evaluation} to convert between \proglang{R} objects
and Protocol Buffer messages.

@@

\begin{verbatim}
  GET https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb
\end{verbatim}
The postfix \code{/pb} in the URL tells the server to send this
object in the form of a Protocol Buffer message.
% Alternative formats include \code{/json}, \code{/csv}, \code{/rds} and others.
If the request
is successful, OpenCPU returns the serialized object with HTTP status
code 200 and HTTP response header \code{Content-Type: application/x-protobuf}.
The latter is the conventional MIME type that formally notifies the client to
interpret the response as a Protocol Buffer.

Because both HTTP and Protocol Buffers have libraries available for many
languages, clients can be implemented in just a few lines of code. Below
is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a Protocol Buffer message from OpenCPU.
In \proglang{R}, we use the HTTP client from
the \code{httr} package \citep{httr}. In this example we
download a data set which is part of the base \proglang{R} distribution, so we can
verify that the object was transferred without loss of information.

@@

This generates Python module called \code{rexp\_pb2.py}, containing
both the descriptor information as well as methods to read and
manipulate the \proglang{R} object message. We use the
HTTP client from the \code{urllib2} module in our example to retrieve the
encoded Protocol Buffer from the remote server then parse and print it
from Python.

@@

The previous example used a simple \code{HTTP GET} method to retrieve
an \proglang{R} object from a remote service (OpenCPU) encoded as a
Protocol Buffer.
In many cases simple \code{HTTP GET} methods are insufficient, and a
more complete RPC system may need to create compact Protocol Buffers
for each request to send to the remote server in addition to parsing
the response Protocol Buffers.

The OpenCPU framework allows us to do arbitrary \proglang{R} function
calls from within any programming language by encoding the arguments
in the request Protocol Buffer.  The following example \proglang{R}
client code performs the remote function call \code{stats::rnorm(n=42,
mean=100)}. The function arguments (in this case \code{n} and
\code{mean}) as well as the return value (a vector with 42 random
numbers) are transferred using Protocol Buffer messages. RPC in OpenCPU
works like the \code{do.call} function in \proglang{R}, hence all
arguments are contained within a list.
